### 2019-07-11 wasp77.tex

\chapter{The case of WASP-77Ab}\label{chap:wasp77}