### 2019-07-12 wasp19.tex
New chapters for three targets

\chapter{The case of WASP-19b}\label{chap:wasp19}

\section{The system}

\section{Transit Parameters and Physical Properties}

\section{Transit Timing Variations}

\section{Limits on an additional perturber}