### 2019-07-12 wasp77.tex

\chapter{The case of WASP-77Ab}\label{chap:wasp77}

\section{The system}

\section{Transit Parameters and Physical Properties}

\section{Transit Timing Variations}

\section{Limits on an additional perturber}